---
title: "Grass_burn_analysis"
author: "Chris Adlam"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
# Load packages
library(tidyverse)
library(emmeans)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(car)
library(ggplot2)
```

# Year 1: effect of planting treatment pre-burn
## Read in data for year 1
```{r}
yr1_data <- read.csv("data/grass_burn_yr1.csv")
yr1_data$plot <- as.factor(yr1_data$plot)
yr1_data$year <- as.factor(yr1_data$year)
yr1_data$date <- as.factor(yr1_data$date)
glimpse(yr1_data)
# Note that blocks are numbered differently in year 1 and year 2, because in year 1 there is no burn treatment, so each "mainplot" is its own block (so there are 20 blocks in year 1 and 10 in year 2). yr1_data has no column for burn or mainplot.
# Remember to check that all variables are read in correctly and set as factor if necessary.
```

## Generating fake data to test code
```{r}
C_values <- rnorm(20, mean = 5, sd = 1)
F_values <- rnorm(20, mean = 9, sd = 1)
R_values <- rnorm(20, mean = 8, sd = 1)

yr1_data <- yr1_data %>% 
  arrange(planting) %>% 
  mutate(Growth = c(C_values, F_values, R_values))

glimpse(yr1_data)
```



## Effect of planting date?
First, may want to see if the different planting dates had an effect. In the data, the date columns is filled with '1' and '2'. '1' is for feb 28th (blocks #1-10), '2' is for mar 7th (blocks #11-20). Here we have the 6 planting*date combinations, with 10 blocks for each date (alternatively, could say 2 replicates of each planting:date combination in 5 blocks).

First let's visualize the planting effect across treatments and dates:
```{r}
ggplot(yr1_data,aes(x=block,y=Growth)) + ylab('Planting mean growth') +
  stat_summary(aes(group=planting,color=planting),fun.y = mean,geom='line') # This calculates the mean of each planting in each block, and connects them with a line

ggplot(yr1_data,aes(x=date,y=Growth)) +
  stat_summary(aes(group=planting,color=planting),fun.y = mean,geom='line') # This calculates the mean of each planting for each date, and connects them with a line
```

Design table if we consider each group of plantings separate blocks (20 blocks total):
| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | planting                      | Categorical | 3        | plot                |
| Treatment | date                          | Categorical | 2        | block               |
| Treatment | planting:date                 | Categorical | 6        | plot                |
| Design    | block                         | Categorical | 20       |                     |
|           | planting:block                | Categorical | 60       |                     |
|           | date:block                    | Categorical | 20       |                     |
|           | date:block:planting           | Categorical | 60       |                     |
|           | plot                          | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |

Design table for evaluating planting date effects if we consider 2 replicates of each combination in 5 blocks:

| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | planting                      | Categorical | 3        | plot                |
| Treatment | date                          | Categorical | 2        | block               |
| Treatment | planting:date                 | Categorical | 6        | plot                |
| Design    | block                         | Categorical | 10       |                     |
|           | planting:block                | Categorical | 30       |                     |
|           | date:block                    | Categorical | 20       |                     |
|           | date:block:planting           | Categorical | 60       |                     |
|           | plot                          | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |

```{r}
# As factorial: plot as EU; field not random. This can tell us the effect of block:planting:planting (?)
date_factorial = lm(Growth ~ date*planting + block,yr1_data)

anova(date_factorial)

date_estimates = emmeans(date_factorial,pairwise~date)
summary(date_estimates$contrasts,infer=T)

# Or analyze as RCBD with random blocks and replicates of treatments in blocks:
date_model = lmer(Growth ~ planting*date + (1|block) + (1|planting:block) + (1|date:block), yr1_data)

anova(date_model)

date_estimates <- emmeans(date_model,pairwise~date,mode='k')
summary(date_estimates$contrasts,infer=T)
```


## Design table year 1: pre-burn treatment
**Design**: RCBD

| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | planting                      | Categorical | 3        | planting:block      |
| Design    | block                         | Categorical | 20       |                     |
|           | planting:block                | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |

## Nurse model
This is to test the effect of the planting treatments (annual grasses and control) on survival and growth of native seedlings during the first year.

```{r}
# Note that the response is not well defined yet.
nurse_model <- lm(Growth ~ planting + block, yr1_data)

# Or, with block as random:
#nurse_model <- lmer(Growth ~ planting + (1|block), yr1_data)

```

## Diagnostic plots
```{r}
# First a regular plot to visualize the data (as above):
ggplot(yr1_data, aes(x = block, y = Growth)) +
  stat_summary(aes(group=planting,color=planting),fun.y = mean,geom='line')

# Diagnostic plots
par(mfrow=c(1,2))
plot(nurse_model,which=2:3)
```

## Anova
```{r}
anova(nurse_model)
```

## Pairwise comparisons
```{r}
planting_estimates <- emmeans(nurse_model,pairwise~planting,mode='k')
summary(planting_estimates$contrasts,infer=T)
```


# Year 2: effect of burn and planting treatments

## Read in data for year 2
```{r}
yr2_data <- read.csv("data/grass_burn_yr2.csv")
yr2_data$subplot <- as.factor(yr2_data$subplot)
yr2_data$block <- as.factor(yr2_data$block)

#need to fix subplot

# Note that blocks are numbered differently in year 1 and year 2, because in year 1 there is no burn treatment, so each "mainplot" is its own block (so there are 20 blocks in year 1 and 10 in year 2).
# Also date of planting is not included in year 2, but now we have mainplot and subplot.
```

## Generating fake data to test code
```{r}
C_values <- rnorm(20, mean = 5, sd = 1)
F_values <- rnorm(20, mean = 9, sd = 1)
R_values <- rnorm(20, mean = 8, sd = 1)

yr2_data <- yr2_data %>% 
  arrange(planting) %>% 
  mutate(Growth = c(C_values, F_values, R_values))

glimpse(yr2_data)
```

## Design table year 2: planting + burn
**Design**: Split plot with RCBD as a main plot

| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | planting                      | Categorical | 3        | planting:block      |
|           | burn                          | Categorical | 2        | burn:block          |
|           | planting:burning              | Categorical | 6        | planting:burn:block |
| Design    | block                         | Categorical | 10       |                     |
|           | MainPlot                      | Categorical | 20       |                     |
|           | SubPlot                       | Categorical | 60       |                     |
|           | planting:block                | Categorical | 30       |                     |
|           | burn:block (~MainPlot)        | Categorical | 20       |                     |
|           | planting:burn:block  (SP?)    | Categorical | 60       |                     |
|           | burn:MainPlot                 | Categorical | 20       |          ?          |
|           | planting:burn:MainPlot (SP?)  | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |


```{r}
# summarize data
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df <- data_summary(yr2_data, varname="LB_num", groupnames="planting")
df
```


## Model
```{r}
# Note that response is not well defined yet.
burn_model = lmer(AM_num ~ planting*burn + block + (1|mainplot) + block:planting, yr2_data[-c(32:61),])

burn_model = lmer(LB_num ~ planting*burn + (1|block) + (1|burn:block) + (1|block:planting), yr2_data[-c(32:61),])
```

## Diagnostic plots
```{r}
# Interaction plot
ggplot(yr2_data[-c(32:61),], aes(x=block,y=LB_num)) + geom_line(aes(group = interaction(planting,burn),color = interaction(planting,burn)))

##Resid v. fitted
plot(burn_model)

## qq plot
qqPlot(resid(burn_model))

## approximate S/L plot
fitted_values = fitted(burn_model)
abs_sqrt_resids = sqrt(abs(resid(burn_model,scaled=T)))
plot(abs_sqrt_resids~fitted_values)
lines(fitted(loess(abs_sqrt_resids~fitted_values)))

## qqplot for random effect
qqPlot(ranef(burn_model)$`mainplot`[,1])
```

## Anova
```{r}
anova(burn_model,ddf='Kenward-Roger')  
```

## Means comparisons
```{r}

# All means estimates
estimates <- emmeans(burn_model,~planting:burn,mode='k')
cld(estimates)

# or:
estimates = emmeans(burn_model,~planting|burn,mode = 'k')
cld(estimates,level = 0.05/3) # adjusted for 4 sets of comparisons (4 planting trts)

# Note that is any of the treatments are not significant, they can be averaged and the data analyzed like a simple RCBD. For example, if burning doesn't have a detectable effect:
## average planting treatments:
planting_data <- aggregate(AM_num~planting+block,yr2_data,mean)
str(planting_data)
## Model with planting treatments averaged:
planting_avg_model <- lm(AM_num~planting+block, planting_data)
anova(planting_avg_model)
## Mean differences:
planting_differences <- emmeans(planting_avg_model,pairwise~planting,mode='k')
summary(planting_differences$contrasts, infer = T)

## Or if planting has no effect:
burn_data <- aggregate(AM_cov~mainplot+burn+block,yr2_data,mean)
str(burn_data)

# Model with burn treatments averaged (straightforward RCBD with block as fixed):
burn_avg_model_fixed <- lm(AM_num~planting + block, planting_data)
anova(burn_avg_model_fixed)

## Mean differences
burn_differences <- emmeans(burn_avg_model_fixed,pairwise~burn,mode='k')
summary(burn_differences$contrasts, infer = T)

## Or just make a model ignoring planting (maybe?):
##subplot_model_fixed = lmer(Growth ~ block + (1|mainplot) + burn + burn:block,yr2_data)
##anova(subplot_model_fixed,ddf='k')
```

## Reporting effects
```{r}
# If there is an interaction between burn and planting, report simple effects:
estimates <- emmeans(burn_model,~burn|planting,mode = 'k')
cld(estimates,level = 0.05/3) # adjusted for 3 sets of comparisons (3 planting treatments)

# or report main effects:
summary(emmeans(burn_model,pairwise~burn,mode='k')$contrasts)
summary(emmeans(burn_model,pairwise~planting,mode='k')$contrasts)

# or comparisons between all combinations:
summary(emmeans(burn_model,pairwise~burn:planting,mode='k')$contrasts)
```

