---
title: "Grass_burn_analysis"
author: "Chris Adlam"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
# Load packages
library(tidyverse)
library(emmeans)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(car)
```

# Read in data for year 1
```{r}
yr1_data <- read.csv("../data/grass_burn_yr1.csv")
```


# Design table year 1: pre-burn treatment
**Design**: RCBD

| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | Planting                      | Categorical | 3        | Planting:Block      |
| Design    | Block                         | Categorical | 20       |                     |
|           | Planting:Block                | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |

# Nurse model
This is to test the effect of the planting treatments (annual grasses and control) on survival and growth of native seedlings during the first year.
```{r}
nurse_model <- lm(Growth ~ Planting + Block, yr1_data)
```


# Read in data for year 1
```{r}
yr2_data <- read.csv("../data/grass_burn_yr2.csv")
```


# Design table year 2: Planting + Burn
**Design**: Split plot with RCBD as a main plot

| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | Planting                      | Categorical | 3        | Planting:Block      |
|           | Burn                          | Categorical | 2        | Burn:Block          |
|           | Planting:Burning              | Categorical | 6        | Planting:Burn:Block |
| Design    | Block                         | Categorical | 10       |                     |
|           | MainPlot                      | Categorical | 20       |                     |
|           | SubPlot                       | Categorical | 60       |                     |
|           | Planting:Block                | Categorical | 30       |                     |
|           | Burn:Block (~MainPlot)        | Categorical | 20       |                     |
|           | Planting:Burn:Block  (SP?)    | Categorical | 60       |                     |
|           | Burn:MainPlot                 | Categorical | 20       |          ?          |
|           | Planting:Burn:MainPlot (SP?)  | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |


# Model
```{r}
# Note that response is not well defined yet.
```

