---
title: "Grass_burn_analysis"
author: "Chris Adlam"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Design**: Split plot with RCBD as a main plot (just RCBD in first year when burn treatment hasn't been applied)

| Structure | Variable                      | Type        | # levels | Experimental Unit   |
|-----------|-------------------------------|-------------|----------|---------------------|
| Treatment | Planting                      | Categorical | 3        | Planting:Block      |
|           | Burn                          | Categorical | 2        | Burn:Block          |
|           | Planting:Burning              | Categorical | 6        | Planting:Burn:Block |
| Design    | Block                         | Categorical | 10       |                     |
|           | MainPlot                      | Categorical | 20       |                     |
|           | SubPlot                       | Categorical | 60       |                     |
|           | Planting:Block                | Categorical | 30       |                     |
|           | Burn:Block                    | Categorical | 20       |                     |
|           | Planting:Burn:Block           | Categorical | 60       |                     |
|           | Burn:MainPlot                 | Categorical | 60       |          ?          |
|           | Planting:Burn:MainPlot        | Categorical | 60       |                     |
| Response  | Growth                        | Numeric     | 60       |                     |
| Response  | Survival                      | Numeric     | 60       |                     |



```{r}

```

